PineAPPL:\\ EW and mixed QCD--EW corrections for PDF fits

### 2020-08-06 paper/app-pineappl.tex
Fix reference

\paragraph{Convolutions.}
The most important subcommand is \texttt{convolute}, which performs a convolution of a single grid with a single or multiple PDF sets.
As an example we show the grid produced for the ATLAS Drell--Yan high-mass lepton-pair production from section~\ref{subsec:numerical_results}, convoluted with \texttt{NNPDF31\_nlo\_as\_0118\_luxqed} as the main PDF set and with \texttt{CT18NLO} as a second PDF set.
\begin{verbatim}
 $ pineappl convolute ATLASZHIGHMASS49FB.pineappl \
 > NNPDF31_nlo_as_0118_luxqed CT18NLO
bin xmin xmax     diff        integ    neg unc pos unc       CT18NLO
---+----+----+------------+-----------+-------+-------+------------+------
  0  116  130 2.0630698e-1  2.888297e0  -2.08%   1.69% 2.0246802e-1 -1.86%
  1  130  150 9.1818985e-2  1.836379e0  -1.79%   1.86% 8.9766355e-2 -2.24%
  2  150  170 4.5306370e-2 9.061274e-1  -1.60%   1.98% 4.4115960e-2 -2.63%
  3  170  190 2.5894856e-2 5.178971e-1  -1.66%   2.06% 2.5138525e-2 -2.92%
  4  190  210 1.6075267e-2 3.215053e-1  -1.70%   2.10% 1.5566535e-2 -3.16%
  5  210  230 1.0526659e-2 2.105331e-1  -1.71%   2.12% 1.0173163e-2 -3.36%
  6  230  250 7.1928162e-3 1.438563e-1  -1.71%   2.13% 6.9403972e-3 -3.51%
  7  250  300 4.0776555e-3 2.038827e-1  -1.70%   2.44% 3.9255068e-3 -3.73%
  8  300  400 1.4775481e-3 1.477548e-1  -1.94%   2.87% 1.4182754e-3 -4.01%
  9  400  500 4.5473785e-4 4.547378e-2  -2.30%   3.19% 4.3525336e-4 -4.28%
 10  500  700 1.2164277e-4 2.432855e-2  -2.41%   3.12% 1.1612523e-4 -4.54%
 11  700 1000 1.9792340e-5 5.937701e-3  -2.05%   2.12% 1.8813113e-5 -4.95%
 12 1000 1500 2.0228761e-6 1.011438e-3  -1.29%   0.47% 1.9221978e-6 -4.98%
\end{verbatim}
The output shows all 13 bins with lower (\texttt{xmin}) and upper limit (\texttt{xmax}) of the invariant mass $M_{\ell \bar{\ell}}$ of the lepton pair, with the differential cross section $\mathrm{d} \sigma / \mathrm{d} M_{\ell \bar{\ell}}$ (\texttt{diff}), integrated cross section $(M_{\ell \bar{\ell}}^\mathrm{max} - M_{\ell \bar{\ell}}^\mathrm{min}) \mathrm{d} \sigma / \mathrm{d} M_{\ell \bar{\ell}}$ (\texttt{integ}), and the perturbative uncertainty estimated from a 7-point scale variation (envelope given by \texttt{neg unc} and \texttt{pos unc}).
The uncertainty estimation can alternatively use a 3- or a 9-point scale variation using the optional program switch \texttt{-{}-scales 3} or \texttt{-{}-scales 9}, respectively.
The (differential) results for the second PDF set (\texttt{CT18NLO}) is shown in absolute numbers and also as a percentage relative to the result of the first PDF set.

@@

The following run card was used to produce the results shown in section~\ref{subsec:numerical_results}.
The only difference with respect to a standard \textsc{mg5\_aMC} run is the switch \texttt{set iappl 1}, which enables to fill a \textsc{PineAPPL} grid.
For a complete set of runcards and patches see TODO.
\begin{verbatim}
set complex_mass_scheme True
import model loop_qcd_qed_sm_Gmu
define p = p b b~
define j = p
generate p p > e+ e- [QCD QED]
output @OUTPUT@
launch @OUTPUT@
fixed_order = ON
set mz @MZ@
set ymt @YMT@
set ebeam1 3500
set ebeam2 3500
set pdlabel lhapdf
set lhaid 324900
set fixed_ren_scale True
set fixed_fac_scale True
set mur_ref_fixed @MZ@
set muf_ref_fixed @MZ@
set reweight_scale True
set ptl = 25.0
set etal = 2.5
set mll = 116
#user_defined_cut set mmllmax = 1500.0
set req_acc_FO 0.0001
set iappl 1
done
quit
\end{verbatim}

>>>>

\paragraph{Convolutions.}
The most important subcommand is \texttt{convolute}, which performs a convolution of a single grid with a single or multiple PDF sets.
As an example we show the grid produced for the ATLAS Drell--Yan high-mass lepton-pair production from section~\ref{sec:dy-lepton-pair-production}, convoluted with \texttt{NNPDF31\_nlo\_as\_0118\_luxqed} as the main PDF set and with \texttt{CT18NLO} as a second PDF set.
\begin{verbatim}
 $ pineappl convolute ATLASZHIGHMASS49FB.pineappl \
 > NNPDF31_nlo_as_0118_luxqed CT18NLO
bin xmin xmax     diff        integ    neg unc pos unc       CT18NLO
---+----+----+------------+-----------+-------+-------+------------+------
  0  116  130 2.0630698e-1  2.888297e0  -2.08%   1.69% 2.0246802e-1 -1.86%
  1  130  150 9.1818985e-2  1.836379e0  -1.79%   1.86% 8.9766355e-2 -2.24%
  2  150  170 4.5306370e-2 9.061274e-1  -1.60%   1.98% 4.4115960e-2 -2.63%
  3  170  190 2.5894856e-2 5.178971e-1  -1.66%   2.06% 2.5138525e-2 -2.92%
  4  190  210 1.6075267e-2 3.215053e-1  -1.70%   2.10% 1.5566535e-2 -3.16%
  5  210  230 1.0526659e-2 2.105331e-1  -1.71%   2.12% 1.0173163e-2 -3.36%
  6  230  250 7.1928162e-3 1.438563e-1  -1.71%   2.13% 6.9403972e-3 -3.51%
  7  250  300 4.0776555e-3 2.038827e-1  -1.70%   2.44% 3.9255068e-3 -3.73%
  8  300  400 1.4775481e-3 1.477548e-1  -1.94%   2.87% 1.4182754e-3 -4.01%
  9  400  500 4.5473785e-4 4.547378e-2  -2.30%   3.19% 4.3525336e-4 -4.28%
 10  500  700 1.2164277e-4 2.432855e-2  -2.41%   3.12% 1.1612523e-4 -4.54%
 11  700 1000 1.9792340e-5 5.937701e-3  -2.05%   2.12% 1.8813113e-5 -4.95%
 12 1000 1500 2.0228761e-6 1.011438e-3  -1.29%   0.47% 1.9221978e-6 -4.98%
\end{verbatim}
The output shows all 13 bins with lower (\texttt{xmin}) and upper limit (\texttt{xmax}) of the invariant mass $M_{\ell \bar{\ell}}$ of the lepton pair, with the differential cross section $\mathrm{d} \sigma / \mathrm{d} M_{\ell \bar{\ell}}$ (\texttt{diff}), integrated cross section $(M_{\ell \bar{\ell}}^\mathrm{max} - M_{\ell \bar{\ell}}^\mathrm{min}) \mathrm{d} \sigma / \mathrm{d} M_{\ell \bar{\ell}}$ (\texttt{integ}), and the perturbative uncertainty estimated from a 7-point scale variation (envelope given by \texttt{neg unc} and \texttt{pos unc}).
The uncertainty estimation can alternatively use a 3- or a 9-point scale variation using the optional program switch \texttt{-{}-scales 3} or \texttt{-{}-scales 9}, respectively.
The (differential) results for the second PDF set (\texttt{CT18NLO}) is shown in absolute numbers and also as a percentage relative to the result of the first PDF set.

@@

The following run card was used to produce the results shown in section~\ref{sec:dy-lepton-pair-production}.
The only difference with respect to a standard \textsc{mg5\_aMC} run is the switch \texttt{set iappl 1}, which enables to fill a \textsc{PineAPPL} grid.
For a complete set of runcards and patches see TODO.
\begin{verbatim}
set complex_mass_scheme True
import model loop_qcd_qed_sm_Gmu
define p = p b b~
define j = p
generate p p > e+ e- [QCD QED]
output @OUTPUT@
launch @OUTPUT@
fixed_order = ON
set mz @MZ@
set ymt @YMT@
set ebeam1 3500
set ebeam2 3500
set pdlabel lhapdf
set lhaid 324900
set fixed_ren_scale True
set fixed_fac_scale True
set mur_ref_fixed @MZ@
set muf_ref_fixed @MZ@
set reweight_scale True
set ptl = 25.0
set etal = 2.5
set mll = 116
#user_defined_cut set mmllmax = 1500.0
set req_acc_FO 0.0001
set iappl 1
done
quit
\end{verbatim}

### 2020-08-06 paper/sec-pineappl.tex
Fix reference

Using 4-tuples has the clear advantage of being very easy to implement and test.
Furthermore, they reproduce the exact numerical value that is also calculated by the MC integrator.
However, the price one has to pay is the size of the 4-tuples.
To give a concrete example, NLO QCD+EW Drell--Yan lepton-pair production (see section~\ref{sec:pineappl-example} and section~\ref{subsec:numerical_results}) needs \SI{159}{\giga\byte} of storage for a target precision of \SI{1}{\percent} of the integrated cross section.
While this is an acceptable size, increasing the precision by an order of magnitude would require roughly 100 times the size, due to the Monte Carlo convergence that goes as $1/\sqrt{N}$ with $N$ being the number of 4-tuples.
With increasing size also the speed of the convolution degrades, because it basically becomes bound by the speed with which the 4-tuples can be read from disk.
However, due to the uncompressed nature of this representation it can serve as an intermediate format to develop and quickly cross check more space-efficient representations, which we will discuss next.

>>>>

Using 4-tuples has the clear advantage of being very easy to implement and test.
Furthermore, they reproduce the exact numerical value that is also calculated by the MC integrator.
However, the price one has to pay is the size of the 4-tuples.
To give a concrete example, NLO QCD+EW Drell--Yan lepton-pair production (see section~\ref{sec:pineappl-example} and section~\ref{sec:dy-lepton-pair-production}) needs \SI{159}{\giga\byte} of storage for a target precision of \SI{1}{\percent} of the integrated cross section.
While this is an acceptable size, increasing the precision by an order of magnitude would require roughly 100 times the size, due to the Monte Carlo convergence that goes as $1/\sqrt{N}$ with $N$ being the number of 4-tuples.
With increasing size also the speed of the convolution degrades, because it basically becomes bound by the speed with which the 4-tuples can be read from disk.
However, due to the uncompressed nature of this representation it can serve as an intermediate format to develop and quickly cross check more space-efficient representations, which we will discuss next.
